---
title: "PA 3: Identify the Mystery College 🏫"
subtitle: "Data Wrangling with `dplyr`"
format: html
code-tools: true
toc: true
editor: source
execute: 
  error: true
  echo: true
  message: false
  warning: false
---

Today you will use the `dplyr` package to clean some data. We will then use that cleaned data to figure out what college Margaret has been accepted to.

::: callout-caution
Make sure you move your *.qmd* file from the Downloads folder into your Stat 331 directory -- you should be working in your R project! **DO NOT** open it straight from download.
:::

# Part One: Data Import and Cleaning

This section will clean today's data so that you can use it more easily in Part Two.

## Data Download & Package Loading

First, we declare our package dependencies and load the data.

::: callout-warning
The data loading function `read_csv()` will give you an outpouring of helpful information about the dataset. If you do not see the word "error", there is nothing to be concerned about.
:::

```{r}
#| label: setup
#| warning: false
#| message: false

library(tidyverse)
colleges <- read_csv("https://www.dropbox.com/s/bt5hvctdevhbq6j/colleges.csv?dl=1")
```

Take a look at the variables in your downloaded data by running the following code. This code with the `str` (structure) function reports the data type for each column in the dataset.

```{r}
str(colleges, give.attr = FALSE)
```


## Data Cleaning

Now we will clean the data. Alas, each of the R chunks in this section will cause an error and / or do the desired task incorrectly. Even the chunks that run without error **are not** correct! You will need to find the mistake and correct it to complete the intended action.

**Step 1:** There are too many variables in this data set. We don't need all of them. Narrow your data set down to only:

-   `INSTNM` name of the institution
-   `CITY` city, `STABBR` state, and `ZIP` ZIP code of the institution
-   `ADM_RATE` admissions rate
-   `SAT_AVG` average SAT score
-   `UGDS` number of undergraduate students
-    `TUITIONFEE_IN` in- and `TUITIONFEE_OUT` out-of-state tuition
-   `CONTROL` Whether the school is public or private
-   `REGION` region of the school.

```{r select-vars}
colleges_clean <- colleges | >  
  select(INSTNM, CITY, STABBR, ZIP,
         ADM_RATE, SAT_AVG, UGDS,
         TUITIONFEE_IN, TUITIONFEE_OUT
         CONTROL, REGION) 
```

**Step 2:** Remove the schools that are for-profit (category 3), keeping public (category 1) and private schools (category 2).

```{r filter-private-profit}
colleges_clean <- colleges_clean |> 
  filter(CONTROL == 1, CONTROL == 2)
```

**Step 3:** Adjust the appropriate variables to be numeric, using `as.numeric()`. 

```{r mutate-numeric}
colleges_clean <- colleges_clean |> 
  mutate(TUITIONFEE_IN  = numeric(TUITIONFEE_IN),
         TUITIONFEE_OUT = numeric(TUITIONFEE_OUT),
         SAT_AVG        = numeric(SAT_AVG),
         UGDS           = numeric(UGDS),
         ADM_RATE       = numeric(ADM_RATE)) 
```

**Step 4:** Adjust the appropriate variables to be factors, using `as.factor()`.

::: callout-note
We will talk more about special data types (including factors) in a few weeks. 
:::

```{r mutate-factor}
colleges_clean <- colleges_clean |>
  mutate(CONTROL = as.character(CONTROL),
         REGION  = as.character(REGION))
```

**Step 5:** Create a new variable called `TUITION_DIFF` which contains the difference between out-of-state and in-state costs.

```{r mutate-differences}
colleges_clean |> 
    TUITION_DIFF = TUITIONFEE_OUT - TUITIONFEE_IN
```

**Step 6:** Create a new variable called `TOTAL_IN` which contains the total amount of money made from tuition per year. (Note, this is just an approximation by mutliplying the number of undergrads by the in-state tuition)

```{r mutate-total}
colleges_clean <- colleges_clean |> 
    select(TOTAL_IN = UGDS x TUITIONFEE_IN)
```

**Step 7:** Remove every row with missing data.

::: callout-warning
This is not always a great idea! Usually, even if *some* of the information is missing, we don't want to throw out the entire row. This time, however, we'll be lazy.
:::

```{r drop-na}
colleges_clean <- colleges_clean |> 
  drop.na()
```

Lastly, notice that each of these steps started with

```{r all-together}
#| eval: false
colleges_clean <- colleges_clean |> ...
```

That is pretty redundant! Instead, we could perform all these tasks as one long "pipeline."

**Step 8:** Combine your (fixed) code chunks into a **single** code chunk that carries out all of the steps necessary to clean the data and save it as `colleges_clean`.

::: callout-tip
Think about coding efficiency -- you **should not** have multiple calls to the **same** function!
:::

```{r combine}
# Code combining ALL of your previous steps into ONE pipeline
```

# Part Two: Identify the Mystery College

Wow! Your best friend Margaret has been accepted to her top choice college! Unfortunately, Margaret is a very mysterious person, and she won't tell you directly which college this is. You'll have to use her clues to figure out which school is her dream school.


You should use the **cleaned version** of the data (`colleges_clean`) and these clues to find her dream school!

**Clues:**

1.  This college is located in Region 4.

2.  This college's admission rate is above the median rate for the region.

3.  This college charges *more* for out-of-state tuition than in-state tuition.

4.  More than 10,000 undergraduates attend this college.

5.  The total amount of money made from tuition per year for this college is at least $200 million. (Use the `TOTAL_IN` variable you created)

6.  The average SAT score of this college is an odd number.

7.  Out of the two remaining colleges, this college has the **least expensive** tuition. While you could just glance at the result to find this, practice using a dplyr verb to get down to one college.

```{r identify-college}
# Code to solve for the mystery college using the 7 clues above.
```

Submit the college Margaret will attend to the Canvas Quiz.
