---
title: "STAT 331 Week 2 Day 2 Handout"
format: html
embed-resources: true
---

```{r}
#| label: setup
#| message: false
#| echo: false

library(tidyverse)
library(RColorBrewer)
```

## Data Import

a. What is a delimiter? What does it do in a data file?


b. Download and unzip the Ages_Data.zip file from Canvas. Save the unzipped directory in a reasonable place, like where you have this Quarto document!

Use the `readr` and `readxl` packages to do the following:

- load the appropriate package(s) if they aren't already
- read in all four data files in the `Ages_Data` directory
- use `read_csv()` to read *ages.csv* with the variable "Name" as a factor data type and "Age" as a character data type.


```{r}
#| label: read-data

```

c. Look over the `readr` and `readxl` cheatsheets. What is at least one functionality that would likely be frequently useful. (This is subjective, I just want you to think about how you can use the packages for data files in the future.)



## Graphics Guidelines

a. Why would you want to use double encoding in a plot?



b. What are some guidelines for using colors in plots?



## Color

We loaded `RColorBrewer` at the beginning of this document.

You can look at pallete options using `display.brewer.all()` or by looking at
[https://colorbrewer2.org/](https://colorbrewer2.org/):

```{r}
display.brewer.all(n = 3, type = "qual" , colorblindFriendly = TRUE)
```

```{r}
mycols <- brewer.pal(n = 3, name = "Set2")
mycols
```

The nicest way to define colors yourself is with their Hex code! The vector `mycols` is a vector of Hex codes.

We can then input this vector of Hex codes as the `value` argument in the `scale_color_manual()` function in `ggplot2`:

```{r}
data(mpg)

ggplot(data = mpg,
       mapping = aes(x = cty,
                     y = hwy,
                     color = drv)) +
  scale_color_manual(values = mycols) + 
  geom_point() +
  facet_grid(cols = vars(year))
```


a. Give it a try yourself! Change the colors in the plot below to an appropriate color scheme.

```{r}
ggplot(data = mpg) + 
  geom_jitter(mapping = aes(x = displ, 
                            y = hwy, 
                            color = cyl)) + 
  labs(x = "Engine Displacement (liters)", 
       y = "Highway MPG", 
       color = "Number of \nCylinders",
       title = "Car Efficiency")
```