---
title: "STAT 331 Week 4 Day 2 Handout"
format: html
embed-resources: true
---

```{r}
#| label: setup
#| message: false
#| echo: false

library(tidyverse)
library(readxl)
library(janitor)
```

## Data

::: callout-warning
For this handout, you need to download TS_data.xlsx from the Week 4 Schedule Canvas page and save it on your computer.
:::

Change the file path below as needed to read in the imdb_data.Rdata file that we used on Monday. 

```{r}
load(file = "data/imdb_data.Rdata")
```

Save the `TS_data.xlsx` file to a reasonable directory, and change the file path below as needed to read it in.

```{r}
eras_data <- read_excel(path = "data/TS_data.xlsx")
```

Change the file path below to read in the military spending data that we used for PA3.

```{r}
military <- read_xlsx("data/SIPRI-Milex-data-1949-2024.xlsx", 
                sheet = "Share of Govt. spending", 
                skip  = 7)
```

## Variable Names in R

a. How do column names need to be referenced in `dplyr` if they start with a number or include spaces?



b. What does the `clean_names()` function from the `janitor` package do?


```{r}
military |> 
  clean_names() |> 
  colnames()
```



## Advanced Joins

a. What were the active years of each director?

```{r}

```

b. What happens for directors that have multiple movies in the join below?

```{r}
directors |> 
  inner_join(movies_directors, 
             join_by(id == director_id))
```

c. When do you need to join on multiple keys?


d. What is the general syntax for joining on multiple keys?


## Working with Categorical Variables (Factors)

a. What are the two general purposes for factor variables (in R)?

b. What function have you used previously to create a factor variable?


### `fct()`

a. The `fct()` function in **forcats** serves the same purpose as what function in base R?

b. How are the levels automatically ordered with the `fct()` function? How is this different from base R?


```{r}
eras_data |> 
  mutate(Album = fct(Album)) |> 
  pull(Album)
```
c. Edit the code above to just output just the levels of the `Album` variable.

d. What does including the `levels = ` argument do?

```{r}
eras_data |> 
  mutate(Album = fct(Album,
                     levels = c("Fearless","Speak Now","Red",
                                "1989", "Reputation","Lover",
                                "Folklore", "Evermore","Midnights"))) |> 
  pull(Album)
```

e. What happens if you forget to provide an album that exists in the data above?



### `fct_recode()`

a. What does this function do? When would you want to use it?


b. What happens to non-specified levels in `fct_recode()`?


c. Change the level "Red" to "Red!" and "Fearless" to "Fearless - Taylors Version" in the `Album` variable:

```{r}

```


### `fct_collapse()`

a. What does this function do?  When would you want to use it?



```{r}
eras_data |> 
  mutate(Genre = fct_collapse(.f= Album,
                       "country pop" = c("Fearless"),
                       "pop rock" = c("Speak Now","Red"),
                       "electropop" = c("1989","Reputation","Lover"),
                       "folk pop" = c("Folklore","Evermore"),
                       "alt-pop" = "Midnights"))
```

b. Why would you want to use `fct_collapse()` rather than `case_when()`? (How is the syntax different?)


c. What happens to levels of the variable `.f = ` that you do not include in `fct_collapse()`?
 
```{r}
eras_data |> 
  mutate(genre = fct_collapse(.f= Album,
                       "country pop" = c( "Fearless"),
                       "pop rock" = c("Speak Now","Red"),
                       "electropop" = c("1989","Reputation","Lover"),
                       "folk pop" = c("Folklore","Evermore")
                       )) |> 
  pull(genre) |> 
  levels()
``` 
 
d. Edit the code below to fix this, so that the Midnights album is categorized as "other" in the `genre` variable.

```{r}
eras_data |> 
  mutate(genre = fct_collapse(.f= Album,
                       "country pop" = c( "Fearless"),
                       "pop rock" = c("Speak Now","Red"),
                       "electropop" = c("1989","Reputation","Lover"),
                       "folk pop" = c("Folklore","Evermore")
                       )) |> 
  pull(genre) |> 
  levels()
``` 
 
 
### `fct_relevel()` 

a. What does this function do?  When would you want to use it?

```{r}
eras_data |> 
  mutate(Album = fct_relevel(.f = Album, 
                             c("Fearless","1989","Taylor Swift",
                               "Speak Now","Red","Midnights","Reputation"))) |>
  pull(Album) |>
  levels()
```

b. What happens with levels that are not specified?



### `fct_infreq()`

a. What does `fct_infreq()` do?  When would you want to use it?


```{r}
eras_data |> 
  ggplot() +
  geom_bar(aes(y = fct_infreq(Album)), 
           fill = "#A5C9A5") +
  theme_minimal() +
  labs(x = "Number of Songs",
       y = "",
       subtitle = "Album",
       title = "Songs Played on the Eras Tour")
```

### `fct_reorder()`

a. What does `fct_reorder()` do?  In what setting would you use it?


```{r}
eras_data |> 
  ggplot(aes(x = Length, 
             y = fct_reorder(.f = Album,
                             .x = Length,
                             .fun = mean), 
             fill = Album)) +
  geom_density_ridges() +
  theme_minimal() +
  theme(legend.position = "none")+
  labs(x = "Song Length (mins)",
       y = "",
       subtitle = "Album",
       title = "Songs Played on the Eras Tour")
```


### `fct_reorder2()`

a. What does `fct_reorder2()` do?  In what setting would you use it?


b. What are the arguments to `fct_reorder2()`?


c. Let's improve the plot that we made way back in week 2 of the number of listings over time for five cities in Texas. Use `fct_reorder2()` to reorder the legend below so that the legend aligns with the order of the lines in the plot.

```{r}
data(txhousing)

txhousing |>
  filter(city %in% c("Dallas","Fort Worth", "Austin",
                     "Houston", "El Paso")) |> 
  ggplot() +
  geom_line(mapping = aes(x = date, 
                          y = listings,
                          color = city))
```
