---
title: "STAT 331 Week 2 Day 1 Handout"
format: html
embed-resources: true
---

```{r}
#| label: setup
#| message: false
#| echo: false

library(tidyverse)
```

## Data

Car data:

```{r}
data(mpg)
```

Texas housing data:
```{r}
data(txhousing)

# taking a subset of cities 
# don't worry about this code - we haven't learned how to do this yet!
sm_tx <- txhousing |>
  filter(city %in% c("Dallas","Fort Worth", "Austin",
                     "Houston", "El Paso"))
```


## `ggplot2` layers

```{r}
ggplot(data = mpg)
```

a. Why does it make sense that the above code produces just a blank plot?

*The ggplot() function with just the data argument initiates a plot, but there is nothing else mapped to the aesthetics of the plot like the axes.*


```{r}
ggplot(data = mpg, 
       aes(x = class, y = hwy))
```

b. What kind of variable is `class`? What about `hwy`? What kind of plot (i.e.) geometries could be appropriate here?

*`class` is categorical and `hwy` is quantitative (continuous). The most typical plot would be side-by-side (or stacked) boxplots.*

c. Fill in the code below to add appropriate layers of geometries:

```{r}
ggplot(data = mpg, 
       aes(x = class, y = hwy)) +
  geom_boxplot() # add boxplot geometry
```

*In the slides, we also added `geom_jitter()` to include the raw data points on the plot. This is typically considered good practice since boxplots only show the 5-number summary and can hide things!* 

```{r}
ggplot(data = mpg, 
       aes(x = class, y = hwy)) +
  geom_jitter() + # add jittered raw data points
  geom_boxplot() # add boxplot geometry
```

*To put the points on top of the boxplots, switch the order of the layers*:

```{r}
ggplot(data = mpg, 
       aes(x = class, y = hwy)) +
  geom_boxplot() + # add boxplot geometry
  geom_jitter() # add jittered raw data points
```


## Practice 

Using the `sm_tx` data, make a plot of median house price over time, distinguishing between different cities.


a. Sketch what this plot could look like. 

*(see slides)*

b. What aesthetics and geometrie(s) will you use? Which variables map to each aesthetic?

*I think it makes sense to use a line plot (`geom_line()`) to track the median house price over time. I will map the `date` variable to the `x` aesthetic, `median` to `y`, and `city` to `color`.*


```{r}
#| label: house-price-plot

ggplot(data = sm_tx,
       aes(x = date,
           y = med,
           color = city)) +
  geom_line()
```


## Aesthetics

a. Below is an example of *global* aesthetics:

```{r}
ggplot(data = sm_tx, 
       mapping = aes(x = date, 
                     y = listings,
                     color = city)) +
  geom_line()
```


b. Below is an example of *local* aesthetics:

```{r}
ggplot(data = sm_tx) +
  geom_line(mapping = aes(x = date, 
                          y = listings,
                          color = city))
```

c. Why do the two chunks above result in the same plot?

*The same mapped aeshtetics are getting passed to `geom_line()`.*


```{r}
ggplot(data = sm_tx) +
  geom_line(mapping = aes(x = date,
                          y = listings,
                          color = city)) +
  geom_hline(yintercept = 20000)
```

d. Why did we need to use local aesthetics in the plot above?


e. Below, color is an example of a ______ aesthetic because 
...

```{r}
ggplot(data = sm_tx,
       mapping = aes(x = listings,
                     y = sales,
                     color = city)) +
  geom_point()
```


f. Below, color is an example of a ______ aesthetic because ...

```{r}
ggplot(data = sm_tx) +
  geom_point(mapping = aes(x = listings,
                           y = sales),
             color = "cornflowerblue")
```

g. What does the warning message indicate for the above two plots??
    
## More About Facets

a.  What function do you use to facet by 1 variable?

b.  What function do you use to facet by 2 variables?

c.  Do facets always need to have the same axis scales? If not, how can you change the scales?
      
## Customization

a. Describe what each element of the plot code below does:

       
```{r}
ggplot(mpg, aes(x = fl, fill = drv)) + 
  geom_bar(position = "fill") +
  scale_fill_manual(values = c("#d73027", "#fdae61", "#4575b4"),
                    labels = c("4-wheel","front","rear")) +
  labs(x = "Fuel Type", 
       y = "Proportion", 
       fill = "Drive Train",
       title = "Fuel Type by Car Drive Train")
```

`ggplot(mpg, aes(x = fl, fill = drv))`:

`geom_bar(position = "fill")`: 

`scale_fill_manual`:

`values = c("#d73027", "#fdae61", "#4575b4")`:

`labels = c("4-wheel","front","rear")`:

`labs(...)`: 

`fill = "Drive Train"`:
