---
title: "STAT 331 Week 2 Day 1 Handout"
format: html
embed-resources: true
---

```{r}
#| label: setup
#| message: false
#| echo: false

library(tidyverse)
```

## Data

Car data:

```{r}
data(mpg)
```

Texas housing data:
```{r}
data(txhousing)

# taking a subset of cities 
# don't worry about this code - we haven't learned how to do this yet!
sm_tx <- txhousing |>
  filter(city %in% c("Dallas","Fort Worth", "Austin",
                     "Houston", "El Paso"))
```


## `ggplot2` layers

```{r}
ggplot(data = mpg)
```

a. Why does it make sense that the above code produces just a blank plot?


```{r}
ggplot(data = mpg, 
       aes(x = class, y = hwy))
```

b. What kind of variable is `class`? What about `hwy`? What kind of plot (i.e.) geometries could be appropriate here?


c. Fill in the code below to add appropriate layers of geometries:

```{r}
ggplot(data = mpg, 
       aes(x = class, y = hwy)) 
```


## Practice 

Using the `sm_tx` data, make a plot of median house price over time, distinguishing between different cities.


a. Sketch what this plot could look like. 

b. What aesthetics and geometrie(s) will you use? Which variables map to each aesthetic?


```{r}
# plot code
```


## Aesthetics

a. Below is an example of ______ aesthetics:

```{r}
ggplot(data = sm_tx, 
       mapping = aes(x = date, 
                     y = listings,
                     color = city)) +
  geom_line()
```


b. Below is an example of ______ aesthetics:

```{r}
ggplot(data = sm_tx) +
  geom_line(mapping = aes(x = date, 
                          y = listings,
                          color = city))
```

c. Why do the two chunks above result in the same plot?


```{r}
ggplot(data = sm_tx) +
  geom_line(mapping = aes(x = date,
                          y = listings,
                          color = city)) +
  geom_hline(yintercept = 20000)
```

d. Why did we need to use local aesthetics in the plot above?


e. Below, color is an example of a ______ aesthetic because 
...

```{r}
ggplot(data = sm_tx,
       mapping = aes(x = listings,
                     y = sales,
                     color = city)) +
  geom_point()
```


f. Below, color is an example of a ______ aesthetic because ...

```{r}
ggplot(data = sm_tx) +
  geom_point(mapping = aes(x = listings,
                           y = sales),
             color = "cornflowerblue")
```

g. What does the warning message indicate for the above two plots??
    
## More About Facets

a.  What function do you use to facet by 1 variable?

b.  What function do you use to facet by 2 variables?

c.  Do facets always need to have the same axis scales? If not, how can you change the scales?
      
## Customization

a. Describe what each element of the plot code below does:

       
```{r}
ggplot(mpg, aes(x = fl, fill = drv)) + 
  geom_bar(position = "fill") +
  scale_fill_manual(values = c("#d73027", "#fdae61", "#4575b4"),
                    labels = c("4-wheel","front","rear")) +
  labs(x = "Fuel Type", 
       y = "Proportion", 
       fill = "Drive Train",
       title = "Fuel Type by Car Drive Train")
```

`ggplot(mpg, aes(x = fl, fill = drv))`:

`geom_bar(position = "fill")`: 

`scale_fill_manual`:

`values = c("#d73027", "#fdae61", "#4575b4")`:

`labels = c("4-wheel","front","rear")`:

`labs(...)`: 

`fill = "Drive Train"`:
