---
title: "PA 2: Using Data Visualization to Find the Penguins"
author: "your names"
format: html
execute: 
  echo: true
  error: true
---

<!-- Insert your setup code chunk here. This is just good practice. -->

## Getting Started

We will be creating visualizations using the `ggplot2` package.

For this activity, we will be exploring the `penguins` data from the `palmerpenguins` package, which has [fantastic documentation with really awesome artwork](https://allisonhorst.github.io/palmerpenguins/). So, you will need to install the **palmerpenguins** package.

`install.packages("palmerpenguins")`

::: {.callout-caution}

## `install.packages()` in the console NOT in your `.qmd` file!

You should type this into your console and NOT include it in a code chunk in your `.qmd` file. Recall that we only have to install a package **once**, but load it each time we open R. Each time you *render* your `.qmd` file, **ALL**the code chunks are run. Therefore, installing a package in a code chunk would cause R to unnecessarily install the package over and over again. Not good.

:::


## Creating a Setup Code Chunk

1.  Insert a code chunk at the beginning of your document (directly under the YAML).
2.  Name the code chunk `setup`.
3.  Use the hashpipe `#|` to specify a code chunk option that prevents any messages (e.g., from loading in packages) from appearing.
4.  Load in the `tidyverse` **or** `ggplot2` package.
5.  Load in the `palmerpenguins` package.


## Dataset: `penguins`

I like to start by **seeing** the dataset I will be working with, so I am going to pull the `penguins` data into my R environment. Do you see it in the top right Environment tab?

```{r}
data(penguins)
```

You may notice that a dataset called `penquins_raw` also loaded. We will ignore this and focus on the `penguins` dataset.

6. Get to know your data. What are the variables and what units are they measured in? What does each row represent?

# Exploring the Penguins Data

## Step 1: Barchart

7. Create a plot of the frequency of each penguin species in the data.

a. Use <https://excalidraw.com/> (or pen and paper, a tablet, etc.) to create a sketch for this plot. Label the aesthetics that will be needed.

b. Use `ggplot2` to create the plot you sketched above.

```{r}
#| label: barchart


```

<!-- Rotate roles! Project manager becomes coder, coder becomes computer, etc. -->

## Step 2: Histogram or Density Curve

8. Use `ggplot2` to plot the distribution of bill lengths for the penguins included in the dataset.

```{r}
#| label: histogram


```

## Step 3: Scatterplot

9. Use `ggplot2` to plot the relationship between the length of a penguin's bill (`bill_length_mm`) and the depth of their bill (`bill_depth_mm`).

```{r}
#| label: scatterplot

```


## Step 4: Add a Trend Line

10. Add a **linear** trend line to the scatterplot you made above! 

<!-- Rotate roles! Project manager becomes coder, coder becomes computer, etc. -->

## Step 5: Adding A Categorical Variable

11. Building off of the plot you made in Step 4, add an aesthetic to differentiate the species of the penguins in the scatterplot by color for both the points *and* the trend line.

```{r}
#| label: scatterplot-with-one-categorical-variable

```

12. Edit your plot in (11) above so that the points are colored by species, but there is only one overall linear trend line.

```{r}
#| label: scatterplot-one-trend

```


13. Building on your code from (11), add the location of the penguins (`island`) to your visualization. There is more than one method to address this, however, one method will more easiliy allow you to address the quesitons below.

```{r}
#| label: scatterplot-with-two-categorical-variables

```

## Canvas Quiz


Working as a team, use the plots you created to answer the following questions on Canvas.

1.  Which species of penguins is represented **least** in the Palmer Penguins data set?

2. Which species had the weakest relationship between bill length and bill depth?

3.  Which species of penguins are found on **every** island?

4.  Which species of penguins are found **only** on Dream Island?

5.  Which species of penguins are found **only** on Biscoe Island?

