---
title: "PA 5.1: Scrambled Message"
subtitle: "`stringr` + Regular Expressions"
format: html
  embed-resources: true
execute: 
  error: true
  echo: true
  message: false
  warning: false
---


```{r setup}
library(tidyverse)
```

## Working with the Secret Message Data

```{r}
message_data <- read_csv("https://github.com/earobinson95/stat331-calpoly/raw/master/practice-activities/data/scrambled_message.txt")
```

If you preivew the data you'll notice this is a **very** simple data frame, with only one column -- `Word`. This column contains strings (or words).

**Since the Word is stored in a data frame, you will need to use some of the dplyr functions you have learned to work with the data.** For example...

 - if you need to find words that have a certain pattern, then you will likely want to pair a `str_XXXX()` function with `filter()`.

- f you need to make changes to the words, then you will likely want to pair a `str_XXXX()` function with `mutate()`.

 - if you want to summarize patterns in the words, then you will likely want to pair a `str_XXXX()` function with `summarize()`.

::: callout-important

In this activity, a "word" is a set of characters with no white space. That is, even though many of elements of the scrambled mess vector are nonsense, and some have punctuation, you can consider each observation to be a "word".

:::

## Warm-up exercises

<!--
The person who has the most pets starts as the Computer (listening to instructions from the Coder and Project Manager)! The person with the second most pets starts as the Project Manager. The coder needs to propose solutions to the task and explain why.
-->

1.  Print out every word in the scrambled message that *starts* with the letter "m". *Hint: What symbol do you use to indicate the beginning of a string?*

2.  Print out every word in the scrambled message that *ends* with a letter "z". *Hint: What symbol do you use to indicate the end of a string?*

3.  Print out the punctuation symbols that are present in the scrambled message. Your output should just be the unique symbols! *Hint: You don't need to type out every punctuation symbol! There is a shortcut way to look for any punctuation symbol.*

<!--
Switch roles! Coder becomes computer, computer becomes project manager, etc.
-->

4. Print out the words that have more than 16 characters. *Hint: This should take two steps!*

5. Print out the longest word in the scrambled message. *Hint: If you already have the length of each word, you can grab the max easily!*




## Decode the Message

Complete the following steps to decode the message.

1.  Remove any spaces before or after each word.

2.  No word should be longer than 16 characters. Drop all extra characters off the end of each word.

<!--
Switch roles! Coder becomes computer, computer becomes project manager, etc.
-->

3.  Every time you see the word "ugh", with any number of h's, followed by a punctuation mark, delete this.

4.  Replace **all** instances of exactly 2 a's with exactly 2 e's.

5.  Replace **all** z's with t's.

6.  Every word that ends in b, change that to a y.

<!--
Switch roles! Coder becomes computer, computer becomes project manager, etc.
-->

7.  Every word that starts with k (or K), change that to a v.

8.  Recombine all your words into a message with a stringr function.


::: callout-note

### Canvas Quiz Submission

What is the name of the movie the quote is from?

:::
