---
title: "Stat 331/531: Statistical Computing with R Syllabus"
subtitle: "Spring 2026"
---

```{r}
#| include: false
#| eval: true
library(tidyverse)

# Create a calendar for your syllabus ----
# Source: http://svmiller.com/blog/2020/08/a-ggplot-calendar-for-your-semester/
# Format: YYYYMMDD

# What are the full dates of the semester?
semester_dates <- seq(ymd(20260330), ymd(20260612), by = 1)

# Weekday(s) of class
class_wdays <- c("Mon", "Wed")

# Breaks
not_here_dates <- c(
  ymd("20260525")
  # Break
  # seq(ymd(20221017), ymd(20221018), by = 1)
  )

# Dates of Exams
exam_dates <- c(ymd(20260506), 
                ymd(20260606)
                )
finals_week <- seq(ymd(20260607), ymd(20260613), by = 1)
project_dates <- ymd(20260605)

# Custom function for treating the first day of the month as the first week 
# of the month up until the first Sunday 
# (unless Sunday was the start of the month)
wom <- function(date) {
    first <- wday(as.Date(paste(year(date), month(date), 1, sep = "-")))
    return((mday(date) + (first - 2)) %/% 7 + 1)
}

# Create a data frame of dates, assign to Cal
Cal <- tibble(date = seq(ymd(20260330), ymd(20260630), by = 1))  |> 
  mutate(mon      = lubridate::month(date, label = T, abbr = F),
         wkdy     = weekdays(date, abbreviate = T),
         wkdy     = fct_relevel(wkdy, "Sun", "Mon", "Tue", "Wed", "Thu","Fri", "Sat"), # make sure Sunday comes first
         semester = date %in% semester_dates, 
         project  = date %in% project_dates,
         exams    = date %in% exam_dates, # is it an exam?
         not_here = date %in% not_here_dates, # is it a day off?
         exam_wk  = date %in% finals_week,
         day      = lubridate::mday(date), 
         week     = wom(date)
         ) |> 
  mutate(category = case_when(project ~ "Project",
                              exams ~ "Exam",
                              not_here ~ "Cal Poly Holiday",
                              wkdy %in% class_wdays & !not_here & !exam_wk & date %in% semester_dates ~ "Class Day",
                           #   date %in% ymd(20250121) ~ "Class Day",
                              semester ~ "Finals",
                              TRUE ~ "NA")
         )


class_cal <- Cal |> 
  mutate(week = case_when(mon == "March"~ 1,
                          .default = week),
         mon = replace_values(mon, "March" ~ "April")
         ) |> 
  ggplot(aes(wkdy, week)) +
  theme_bw() +
  theme(aspect.ratio = 1, 
        panel.grid.major.x = element_blank(),
        legend.position = "right",
        legend.title = element_blank(),
        axis.title.y = element_blank(), 
        axis.title.x = element_blank(),
        axis.ticks.y = element_blank(), 
        axis.text.y = element_blank()) +
  geom_tile(alpha = 0.8, 
            aes(fill = category), 
            color="black", 
            linewidth = 0.45
            ) +
  facet_wrap(~ mon, scales = "free", ncol=3) +
  geom_text(aes(label = day, color = semester & (!not_here))) +
  # put your y-axis down, flip it, and reverse it
  scale_y_reverse(breaks = NULL) +
  # manually fill scale colors to something you like...
  scale_color_manual(values = c("FALSE" = "grey80", "TRUE" = "black"), guide = "none") + 
  scale_fill_manual(values=c("Class Day"="steelblue", 
                             "Finals"="white",
                             "Cal Poly Holiday" = "grey10",
                             "NA" = "white", # I like these whited out...
                             "Exam"="orange3",
                             "Project"="#c994c7"),
                    #... but also suppress a label for a non-class semester day
                    breaks=c("Finals", "Cal Poly Holiday", "Class Day","Exam","Project"))
# class_cal

exam_days <- Cal |> 
  filter(category == "Exam") |> 
  mutate(topic = c("Midterm Exam", 
                   "Tentative Common Final Exam"
                   )
         ) |> 
  rowwise() |> 
  mutate(sem_week = isoweek(date),
         Date = paste(format(date, "%b %e"), sep = ",", collapse = ", "))

project_days <- Cal |> 
  filter(category == "Project") |> 
  mutate(topic = c("Project Final Report Due")
         ) |> 
  rowwise() |> 
  mutate(sem_week = isoweek(date),
         Date = paste(format(date, "%b %e"), sep = ",", collapse = ", "))



class_days <- Cal |> 
  filter(date %in% semester_dates,
         category %in% c("Class Day")
         ) |>  
  mutate(sem_week = isoweek(date)) |> 
  group_by(sem_week) |> 
  summarize(Date = paste(format(date, "%b %e"), sep = ",", collapse = ", ")) |> 
  ungroup() |>
  mutate(topic = c(
    "Introduction to R",
    "Basics of Graphics",
    "Data Cleaning and Manipulation",
    "Data Transformations + Factors",
    "Special Data Types",
    "Version Control",
    "Functions",
    "Functional Programming",
    "Simulation",
    "Inference and Advanced Graphics")) |> 
  bind_rows(project_days) |> 
  bind_rows(exam_days) |> 
  arrange(sem_week) |> 
  select(Date, Topic = topic)


# Weekday(s) of office hours
officehr_wdays <- c("Tue", "Thu")

# Create a data frame of dates, assign to Cal
Cal2 <- tibble(date = seq(ymd(20240107), ymd(20240113), by = 1))  |> 
  mutate(mon      = lubridate::month(date, label = T, abbr = F),
         wkdy     = weekdays(date, abbreviate = T),
         wkdy     = fct_relevel(wkdy, "Sun", "Mon", "Tue", "Wed", "Thu","Fri", "Sat"), # make sure Sunday comes first
         week     = wom(date),
         classday = wkdy %in% class_wdays,
         officehrday = wkdy %in% officehr_wdays,
         officehrtime = c(NA, NA, "10:30am-\n12:00pm", NA, "9:00-\n10:00am",  NA, NA)
  )


classday_cal <- Cal2 |> 
  ggplot(aes(wkdy, week)) +
  theme_bw() +
  theme(aspect.ratio = 0.15, 
        panel.grid.major.x = element_blank(),
        legend.position = "right",
        legend.title = element_blank(),
        axis.title.y = element_blank(), 
        axis.title.x = element_blank(),
        axis.ticks.y = element_blank(), 
        axis.text.y = element_blank()) +
  geom_tile(alpha = 0.8, 
            aes(fill = classday), 
            color="black", 
            linewidth = 0.45,
            show.legend = F
  ) +
  scale_x_discrete(expand = c(0,0)) +
  scale_y_discrete(expand = c(0,0)) +
  scale_fill_manual(values=c("white", "steelblue")
                    )
# classday_cal

officehr_cal <- Cal2 |> 
  ggplot(aes(wkdy, week)) +
  theme_bw() +
  theme(aspect.ratio = 0.15, 
        panel.grid.major.x = element_blank(),
        legend.position = "right",
        legend.title = element_blank(),
        axis.title.y = element_blank(), 
        axis.title.x = element_blank(),
        axis.ticks.y = element_blank(), 
        axis.text.y = element_blank()) +
  geom_tile(alpha = 0., 
            aes(fill = officehrday), 
            color="black", 
            linewidth = 0.45,
            show.legend = F
  ) +
  geom_text(aes(label = officehrtime), color = "black") +
  scale_x_discrete(expand = c(0,0)) +
  scale_y_discrete(expand = c(0,0)) +
  scale_fill_manual(values=c("white", "gray")
  )
# officehr_cal

```

## Communication

**Dr. Charlotte Zilber Mann**

You can call me Professor, Professor/Prof. C. or Dr. C.

`r fontawesome::fa("envelope")` Email:
[czmann\@calpoly.edu](mailto:czmann@calpoly.edu?subject=Stat%20331/531)

`r fontawesome::fa("building")` Office: Building 25 Office 201

**Email:**

-   Only email me with your @calpoly.edu email – I cannot respond to
    other email addresses
-   Start your email subject line with “\[STAT 331\]”
-   I will do my best to respond within 24 hours during the week
-   If you send me an email past 6pm or on the weekend, I will likely
    not respond until the morning of the next working day
-   *You should only email me about things that relate to you as an
    individual.* Any other questions should be posted on the course
    Discord (more details below).

**Course Discord Page:**

For questions of general interest, such as course clarifications or
conceptual questions, please use the course [Discord page](discord-instructions.html) (you will join
this Week 1). If you have a question – someone else in class does too! I encourage you all to respond to each other. I will do my best to check Discord twice each weekday during working hours (8-6) to ensure that questions are being answered accurately.

I encourage you to give your post a concise and informative initial
sentence, so that other people can find it. For example, *“How do I color bars in a barplot with ggplot?”* is a better opening sentence than *“help with plotting”*.

You may post snippets of your code and errors on Discord, but please do
not post full solutions to lab assignments.

`r fontawesome::fa("brain")` While your posts are not anonymous, in this case there is no such thing as a bad question!

## Course Logistics

**Class Meeting Times:** Mondays / Wednesdays 

```{r}
#| eval: false
#| include: false
#| echo: false
#| warning: false
#| message: false
#| fig-width: 7
#| fig-height: 1.5
#| out-width: 75%
#| fig-align: center
#| fig-pos: H
classday_cal
```

-   Section 70: 8:10am - 10:00am
-   Section 71: 10:10am - 12:00am

**Classroom** 38-122

**Office Hours** are held in my office (25-201) during the following
times:

<!-- ::: column-margin -->

<!-- {{< video https://www.youtube.com/embed/yQq1-_ujXrM?si=sFpcpfT9Cey2-Eaf >}} -->

<!-- Introducing Faculty Office Hours -->

<!-- ::: -->

```{r}
#| echo: false
#| warning: false
#| message: false
#| fig-width: 7
#| fig-height: 1.5
#| out-width: 90%
#| fig-align: center
#| fig-pos: H
officehr_cal
```

<!-- | Day        | Time                | -->

<!-- |------------|---------------------| -->

<!-- | Mondays    | 3:10pm - 4:00pm     | -->

<!-- | Wednesdays | 1:10pm - 2:00pm     | -->

<!-- | Thursdays  | 2:30pm - 3:30pm   | -->

- **Office hours are for you!** During this time, I am available to talk about anything related to the course. 
- **Office hours are also helpful for me…** this is my chance to get to know you all better and also get a good idea of what is most challenging and most interesting to you.
- See [this resource](https://lsc.cornell.edu/consult-with/how-to-use-office-hours/) for ways that students use office hours. 
- One thing I will not do during office hours is tell you whether an answer to a problem is “correct” or not. We can talk through your thinking that led to your solution, but will leave the final assessment to once you have submitted an assignment.
- Note that office hours are not just for when you have content questions.
Stop by to introduce yourself, ask questions about the broader field of
statistics, or share what you are working on!

## Course Description

Stat 335/531 provides you with an introduction to programming for data
and statistical analysis. The course covers basic programming concepts
necessary for statistics, good computing practice, and use of built-in
functions to complete basic statistical analyses.

**Prerequisites**

Entrance to STAT 331/531 requires successful completion of:

-   a Stat II qualifying course, and
-   an introductory programming course.

## Learning Objectives

This course will teach you the foundations of statistical computing
principles in the language of `R`.

After taking this course, you will be able to:

-   Work with the RStudio Integrated development environment (IDE) and
    quarto documents.
-   Import, manage, and clean data from a wide variety of data sources.
-   Visualize and summarize data for informative exploratory data
    analysis and presentations.
-   Write efficient, well-documented, and tidy `R` code.
-   Program random experiments and simulations from probability models.

Additionally, it is my hope that you will learn to:

-   Extend your `R` skills independently through documentation and
    online resources.
-   Be thoughtful, deliberate, and ethical in your use of `R` and
    similar tools.
-   Use `R` to be playful, creative, and fun!

## Course Materials

**Course Webpage (Canvas)**

All course material will be made available on our Canvas page, including check-ins, practice
activities, lab assignments, challenges, review questions, and selected solutions. Canvas will
also contain detailed weekly schedules throughout the term. You are responsible for
checking the Canvas page on a *regular* basis and making sure you see Canvas Announcements.

I will also be posting material on [this website](https://manncz.github.io/stat-331-s26/) for easy access to materials. Please note that **this website does not include deadlines** and links to all material will additionally be posted to Canvas. Canvas will always contain the most updated information about course deadlines and details, so always defer to these. 

**Textbook**

`r fontawesome::fa("book")` [R for Data Science (2nd edition)](https://r4ds.hadley.nz/).

There is an abundance of free online resources for learning programming and R. We will primarily use R for Data Science (2nd edition), by Hadley Wickham, Mine Çetinkaya-Rundel, and Garrett Grolemund for readings.

Each week you can reference readings with some additional commentary and videos through the [schedule](../index.qmd). This material is based on work by [Dr. Emily Robinson](https://github.com/earobinson95), [Dr. Susan
VanderPlas](https://github.com/srvanderplas), [Dr. Allison
Theobold](https://github.com/atheobold), and [Dr. Kelly
Bodwin](https://github.com/kbodwin).

In addition, you may find it useful to reference some of the following
resources during the course. Most are available online for free.

-   [Modern Dive](https://moderndive.com/1-getting-started.html)
-   [Introduction to Modern
    Statistics](https://openintro-ims.netlify.app/)
-   [Advanced R](http://adv-r.had.co.nz/)

**Computing**

`r fontawesome::fa("computer")` You will need access to a laptop for this course. If this requirement is limiting for you, please
contact me ASAP.

We will (obviously) be using the `R` statistical software
throughout this course. In addition, you are required to use
`RStudio`, a companion integrated developer environment (IDE).
Both `R` and `RStudio` are freely available. Download instructions are available on Canvas.

::: callout-warning
Chromebooks and iPads will **not** be
sufficient to use `R`. If this requirement is limiting for you, please
contact me ASAP.
:::


\newpage

## Class Schedule & Topic Outline

This schedule is tentative and subject to change.

```{r calendar}
#| echo: false
#| eval: true
#| warning: false
#| fig-width: 8
#| fig-height: 3.5
#| fig-pos: 'h'
#| fig-cap: "Note that we do not have class on Monday May, 25th for the Memorial Day Holiday."
class_cal
```

```{r schedule}
#| echo: false
#| eval: true
#| fig-pos: 'h'
class_days |> 
kableExtra::kable(caption = "Tentative schedule of class topics and important due dates", format = "simple")
```


## Assessments

### General Evaluation Criteria

In every assignment, discussion, and written component of this class,
you are expected to demonstrate that you are intellectually engaging
with the material. I will evaluate you based on this engagement, which
means that technically correct but low effort answers which do not
demonstrate engagement or understanding will receive no credit.

When you answer questions in this class, your goal is to show that you
either understand the material or are actively engaging with it. If you
did not achieve this goal, then your answer is incomplete, regardless of
whether or not it is technically correct. This is not to encourage you
to add unnecessary complexity to your answer - simple, elegant solutions
are always preferable to unwieldly, complex solutions that accomplish
the same task.

While this is not an English class, grammar and spelling are important,
as is your ability to communicate technical information in writing; both
of these criteria will be used in addition to assignment-specific
rubrics to evaluate your work.

### Formative Assessments

#### Group Quizzes

At the beginning of class on Mondays, you will take a short quiz **on paper** with your working group to check that you have a good understanding of the material from the *previous week*. You will have access to relvant code cheatsheets during the quiz. 

You need to be present with your group to receive credit on the quiz. However, you can have one excused quiz during the quarter.

-   Practice Activities are typically **Mondays at the beginning of class**


#### Practice Activities

Most Tuesdays's, you will be given a Practice Activity to complete *with your group*, to get the hang of the week’s necessary `R` skills. These activities will
always result in a single, straightforward correct answer, that you will
submit via Canvas (two attempts, the average score being kept).

Since these activities are intended to be your first attempt at new
skills, *they are meant to be done with help from me and your peers*.
Therefore, you will always be given some time in class to work on them.

-   Practice Activities are (typically) due **Tuesdays at 11:59pm**


#### Lab Assignments

Your typical homework assignments will be weekly labs. You will follow
each lab's instructions to complete tasks in `R` and submit a rendered
.html Quarto document to Canvas.

Most weeks, there will be class time on Wednesdays dedicated to working
on completing lab assignments. While you will work with peers during
class time on Wednesdays, you will be expected to submit your own
individual work for the final submission.

-   Labs are (typically) due on **Sundays at 11:59pm**.

### Evaluative Assessments

#### Final Project

I am changing the final project from previous quarters, so it is still TBD.

#### Exams

There will be a **midterm exam** and a **final exam**. The midterm will have both in-class and take-home portions. The final will be entirely in-class. I am requesting a Common Final Exam time. More information will follow.


### Course Grade

Your grade in STAT 331/531 will contain the following components:

| Assignments         | Weight |
|---------------------|--------|
| Group Quizzes & Misc.      | 10%    |
| Practice Activities | 10%    |
| Lab Assignments     | 20%    |
| Midterm Exam        | 20%    |
| Final Project       | 15%    |
| Final Exam          | 25%    |

Lower bounds for grade cutoffs are shown in the following table. It is possible that I could lower any of the grade cutoffs. You will only have the opportunity for grade rounding to the nearest tenth of a percent if you completed all assignments in the course. 

| Letter grade | X +  | X   | X - |
|--------------|------|-----|-----|
| A            | .    | 93  | 90  |
| B            | 87   | 83  | 80  |
| C            | 77   | 73  | 70  |
| D            | 67   | 63  | 60  |
| F            | \<60 |     |     |

Interpretation of this table:

-   A grade of 85 will receive a B.
-   A grade of 77 will receive a C+.
-   A grade of 70 will receive a C-.
-   Anything below a 60 will receive an F.

## Course Policies

### Attendance & Participation

I do not take formal attendance in this class. The only time your attendance is required is for the Group Quizzes on Mondays. However, it is my expectation that you remain in class and on task until you have finished all your activities and assignments.

**If you are feeling ill, please do not come to class**.

Here’s what you should do if you do miss a class:

-   Talk to a classmate to figure out what information you missed
-   Check Canvas for any necessary handouts or changes to assignments
-   Post on Discord with any questions you have after reviewing notes
    and handouts

If you miss a bunch of classes, please come talk to me. I’m working from the assumption that you care and are trying, but something is getting in
your way (health issues? college stress? work?); let’s figure out what
that is and how I can help.

### Accessibility and Accomodations

I have taken numerous steps to ensure that all of the materials presented in this course are accessible to all participants, regardless of physical or learning disabilities. I know that everyone is unique and I may have unintentionally overlooked something that limits access to some materials or activities. Please let me know if you cannot access any content. If you find you need additional accommodations to complete the required course work, please contact me as soon as possible! I want to support your success. 

If you wish to request disability-related accommodations such as extra exam time for this or any other course, please contact the [Disability Resource Center](https://drc.calpoly.edu/). It's important to do this with as much advanced notice as possible, so you have full access to your course materials and activities in a timely manner. **Please also email me** to let me know that I should expect an email from the DRC to get you the accommodations that you need.


### Inclusive Classroom 
Because data are collected by and about humans, data necessarily
encode aspects of our proclivities and biases. As a result, this course may touch upon
difficult topics related to race, gender, inequality, class, and oppression. We each come into
this class with different perspectives that can be shared to enhance our understanding of
these issues. I ask that you enter these conversations with respect, curiosity, and cultural
humility. You should be open to alternative perspectives and be willing to revise beliefs that
are based on misinformation. As a general rule, your ideas and experiences can always be
shared during these conversations, but please refrain from dismissing the experiences of
others. Personal attacks of any kind will not be tolerated.

### Group Work

This course will rely heavily on in-class group work. You will be expected to work with classmates on in-class material. If you have concerns about random group assignment, please
talk to me at the start of the term.

You are
also encouraged to discuss your ideas for lab assignments; however, I expect that these
collaborations are about ideas and no R code is shared. Each person’s lab assignment
submissions are expected to reflect their own thinking, and thus copying the work of others
does not provide me with any information about your learning.


The community standards for student-to-student and student-to-instructor interactions include the following:

- Listen actively and attentively.
- Ask for clarification if you are confused.
- Challenge one another respectfully.
- Gracefully accept constructive criticism.
- Take responsibility for the quality of the discussion and the work.
- Do not monopolize discussion.
- Acknowledge that everyone has something to offer.
- Speak from your own experience, without generalizing.


### Late Policy


**Everyone will be given a “bank” of 4 *deadline extensions*** which can be used on Practice Activities and Lab Assignments at any point in the quarter. **One deadline extension will get you a 24 hour extension on a given assignment.** You can use all of your deadline extensions on one assignment or split them up in any way. 

-	Like how the “real world” works, you must request a deadline extension **before the deadline.** To use your deadline extensions, you need to email me **before the assignment deadline.** You don't need to get a response from me, you just need to email me at least a minute before the deadline. 
-	If you do not email me before the assignment deadline (or you run out of late days), you will be docked 20% for each day that the assignment is late. (e.g. you can get a maximum of 80% on the assignment if it is submitted within 24 hours after the original due date and time.)

A note on this policy: my priority is for all of you to engage with and
understand the course materials. The best way to achieve this is to
complete all assignments and receive feedback on your work. At the same
time, I understand that things happen outside of our control and that
you are balancing many responsibilities.

::: callout-warning
### Automatic Canvas Settings

Canvas is set up to automatically input 0% for missing assignments (as
an incentive to go complete the assignment) and apply the 20% grade
deduction policy. I will need to manually adjust your grade when you use
your "deadline extensions." You are responsible for double checking your
grade to make sure I didn't miss anything.
:::

If you find yourself with extenuating circumstances beyond the defined
late policy, please email me before the due date.


### Exam Conflicts

If you have a known conflict with an exam, please discuss it with me **at least three weeks prior to the exam date**. If an illness or a family emergency arises, please
let me know as soon as possible, and we will work out a solution. This may involve taking the
exam at a later date or replacing the missed exam score with the final exam score. If you are
unable to take the final exam, you will likely receive an Incomplete in the course.

## Course Expectations


**For Students, I expect:**

- You will ask lots of questions.
- You will do reading and other assignments outside of class and be prepared for each class meeting.
- You will engage in some productive struggle/challenge when working on assignments (i.e. not give up and google or use ChatGPT immediately!)
- If you find yourself struggling in general or feel that you are falling behind, please reach out to me as soon as possible so that we can develop a plan for your success together. 
- You will take advantage of the resources that I provide and to seek out additional resources if you find that you are struggling with the material. You do not need to seek out additional resources all on your own – we can figure out what additional resources you may need.  
- You will be doing your best to engage with what we are all doing together during class. This means that you are using technology (i.e. laptops/tablets/phones) for class activities and not in ways that distract you. 
- You will approach the course with academic integrity.
- You will come to class on time and work on class material with your peers.
- Everyone can learn the material in this course.

**For Me:**

- I am responsible for ensuring that all students feel welcome and valued in the classroom.
- I am responsible for providing the resources necessary so that all students in the course can achieve the learning objectives.
- I am responsible for providing prompt and clear feedback on your coursework.
- I welcome and appreciate any actionable feedback regarding how the class is meeting your learning needs. We can discuss ways to adjust the course to ensure it is supporting your learning and do more of anything that is working well!



### Make Mistakes!

Programming is the process of making a series of silly or stupid
mistakes, and then slowly fixing each mistake (while adding a few more).
The only way to know how to fix these mistakes (and avoid them in the
future) is to make them. (Sometimes, you have to make the same mistake a
few dozen times before you can avoid it in the future). At some point
during the class, you will find that you've spent 30 minutes staring at
an error caused by a typo, a space, a parenthesis in the wrong place.
You may ask for help debugging this weird error, only to have someone
immediately point out the problem... it is always easier to see these
things in someone else's code. This is part of programming, it is
normal, and you shouldn't feel embarrassed or sorry (unless you put no
effort into troubleshooting the problem before you asked for help)

If you manage to produce an error I haven't seen before, then
congratulations. You have achieved something special, and that
achievement should be celebrated. Each new and bizarre error is an
opportunity to learn a bit more about the programming language, the
operating system, or the interaction between the two.


## Academic Integrity and Class Conduct

**Our academic community is at its best when we treat each other with fairness, honesty, respect, and trust. Unfortunately, sometimes students slip up and do something that gives themselves (or someone else) an unfair advantage over other students.  Such actions will not be tolerated in this class.**

It is most likely that a student will slip up if they are feeling overwhelmed or unsure of how to approach a problem that they are stuck on. So **let's be proactive to prevent those situations!!** If you are helpless in an assignment or in studying, post on discord, come to office hours, or email me or one of your classmates! I am working hard to provide all of the resources you need to succeed


However, **the college doesn't really care if you break the rules on purpose or by accident.** If I suspect you have done something that violates academic integrity on any graded elements of the course (included those graded for completion), you will receive an email to tell you about my suspicion and how the situation will be handled. **Typically, students earn a 0 for the  assignment on which academic misconduct is found.** University policy dictates that we must report every instance of suspected academic dishonesty to the Office of Student Rights and Responsibilities, no matter how small. For more
information on academic misconduct and what constitutes cheating and
plagiarism, please see
[academicprograms.calpoly.edu/content/academicpolicies/Cheating](https://academicprograms.calpoly.edu/content/academicpolicies/Cheating).

It is important to note that paraphrasing or quoting another's work
without proper citation is a form of academic misconduct. This includes
the `R` code produced by someone else! Writing code is like writing a
paper, it is obvious if you copied-and-pasted a sentence from someone
else into your paper because the way each person writes is different.

Even inadvertent or unintentional misuse or appropriation of another’s
work (such as relying heavily on source material that is not expressly
acknowledged) is considered plagiarism. This includes using **Generative AI, which should only be used to help you problem solve just as a tutor or peer would, and not as a substitute for your own work.** If you are
struggling with writing the `R` code for an assignment, please reach out
to me and post to Discord. I would prefer that I get to help you rather than you spending
hours Googling things and get nowhere!

### Generative AI Policy

Generative AI (like Chat GPT) is a powerful tool to help write code. However, it should only be used *once you understand the software*, which is what you are in this class to do!! At this stage of learning R software, your own thoughts and creativity are what you need.

Class time is structured around active learning and I expect you to actively engage with the software, so **you may not use generative AI for in-class work and activities**. You have each other and me to grapple with the material during class.  If you use generative AI in class, there will be a deduction in your "Group Quizzes & Misc." portion of the grade.

You can use generative AI to complete lab assignments. However, I encourage you to give all problems a full effort on your own, referencing the provided course materials and cheatsheets **before** turning to AI. You will learn SO much more this way!! You should also give thought to how you are using AI, and try to approach it more as a tutor, asking it specific questions to help you solve a problem, rather than just feeding it the lab assignment. When you do use AI to help with a problem, make sure you understand **why** the code it provides works. If you choose to use AI, you **must** cite your use including the prompt that you gave.

Note that work created by AI tools may not be considered original work and instead, considered automated plagiarism. It is derived from previously created texts from other sources that the models were trained on, yet doesn't cite sources. Thus, there is a high risk that you can unintentionally plagiarize other sources from the web.


A final personal opinion is that I find YOU so much more interesting than AI! I would so much rather see and read what you are thinking and how you structure code for a problem than something you may think looks more “clean” from AI. 




::: callout-tip
## Always Cite Outside Resources!

In this class the assumed knowledge is the course materials, including the course textbook, coursework pages, and course slides.  If you use outside resources on assignments I expect that you will *cite those resources.* This means that

- If you use any **`R` functions or code that are not in the course materials**, you must cite where you found it. e.g., if you look up the help file for `mutate()` and you learn about a function not covered in the course materials, you are required to cite the documentation when using that function.
- If you use **Google** you are expected to “inform” me of any resources you used by pasting the link to the resource in a code comment next to where you used that resource.
- If you use **AI** you are expected to “inform” me by stating that you used it in a code comment next to where you used it including **the full prompt that you gave AI (like ChatGPT)**.
- If you **work with classmates** on an assignment, you are expected to “inform” me by including a note with their names. 
:::

**If you have any questions about using and citing sources, you are
expected to ask for clarification.**

## Learning Environment and Support

My priority is your success in this course. I understand that things happen unexpectedly and obtaining official accommodations can be slow. 

If you feel that any circumstances or factors beyond your control may affect your presence and work in this class, **please reach out to me** and we will work to accommodate your needs. You do not need to share personal information with me when you reach out, unless you would like to.


### Mental Health

National surveys of college students have consistently found that stress, sleep problems, anxiety, depression, interpersonal concerns, death of a significant other and alcohol use are among the top ten health impediments to academic performance. If you are experiencing any mental health issues, I and Cal Poly are here to help you. Cal Poly’s Counseling Services (805-756-2511) is a free and confidential resource for assistance, support and advocacy [https://chw.calpoly.edu/counseling](https://chw.calpoly.edu/counseling).

### Support Services

If you are experiencing food insecurity, housing instability, or other
challenges that may impact your ability to succeed in this course, please refer
to the resources listed on Canvas under "Student Support Services at Cal Poly."
These resources provide a range of essential support services, including
emergency financial assistance, counseling, and academic support.