---
title: "Week 8 Day 2: Simulation"
number-sections: true
number-depth: 2
---

```{r}
#| echo: false
#| label: setup
#| message: false

library(tidyverse)
library(knitr)
library(palmerpenguins)
data(penguins)
```


# Learning Objectives

This week we will learn about simulating data and running simulations in R. By the end of the week you should:

- know functions for simulating some of the common statistical distributions (e.g., Normal, t, Chi-Square, Binomial)
- know what each of the `r`, `p`, `d`, and `q` functions do when simulating data from different statistical distributions
- know how to plot an empirical and theoretical distribution
- know how to use random sampling with replacement or not in R
- know how to implement multiple iterations of a simulation using a `map` or `pmap` function.

#### 📖 Required Readings: 60 min

------------------------------------------------------------------------

# Simulation in R

Simulation is a very important tool for statistical computing! "Simulation" is a very broad term that encompasses a lot of different things. For our purposes, we will focus on:

- simulating data by drawing from probability distributions
- simulating random processes to get an idea of their behavior

# Working with Common Statistical Distributions in R

#### 📖 Required Reading: [R Programming for Data Science : 20.1 - 20.3](https://bookdown.org/rdpeng/rprogdatascience/simulation.html)

#### 📽 Optional Video: [Introduction to Simulating Data](https://www.youtube.com/watch?v=J5XQaxmznNE)

## Reproducibility -- Setting seeds

Whenever you are working with an operation in R that involves randomness (like generating random samples from a probability distribution!), you  should choose your own seed using the `set.seed()` function. This
guarantees your results will be consistent across runs (and hopefully  computers):

```{r}
set.seed(1234)

rnorm(1, mean = 0, sd = 1)
```

```{r}
set.seed(1234)

rnorm(1, mean = 0, sd = 1)
```

Of course, it doesn't mean the results will be the same in every subsequent run
if you forget or reset the seed in between each line of code!

```{r}
set.seed(1234)

rnorm(1, mean = 0, sd = 1)

## Calling rnorm() again without a seed "resets" the seed! 
rnorm(1, mean = 0, sd = 1)
```

It is very important to always set a seed at the **beginning** of a Quarto
document that contains any random steps, so that your rendered results are
consistent.

Note that you can choose your favorite number to be your seed - any number will do!

::: {.callout-caution}
Note, though, that this only guarantees your rendered results will be the same
*if the code has not changed*.

Changing any part of the code will re-randomize everything that comes after it!
:::

## Reproducibility -- inline code

When writing up a report which includes results from a random generation
process, in order to ensure reproducibility in your document, use `` `{r}` `` or `` `r` `` to
include your output within your written description with [inline code](https://quarto.org/docs/computations/inline-code.html).

::: {.callout-note collapse=TRUE}
# Reproducibility: inline code example

```{r}
#| eval: true
my_rand <- rnorm(1, mean = 0, sd = 1)
my_rand
```

Using `{r} knitr::inline_expr('my_rand', syntax = "md")` will display the result
within my text: 

My random number is `{r} my_rand`.

Alternatively, you could have put the `rnorm` code directly into the inline text
`{r} knitr::inline_expr('rnorm(1, mean = 0, sd = 1)', syntax = "md")`, but this
can get messy if you have a result that requires a larger chunk of code.

::: 

## Plotting Distributions

If you took an intro statistics class at Cal Poly, you likely generated null distributions using both simulation and with a theoretical distribution. It is often helpful to be able to plot both a simulated (or empirical) distribution and a theoretical distribution together to see how well they match.

Let's walk through the example in the lecture slides. 

First, I will create simulated data with 1,000 observations, drawn from a normal distribution with mean 67 and standard deviation 3 ($N(67, 3)$).

```{r}
set.seed(395)

my_samples <- set.seed(93401)
my_samples <- tibble(height = rnorm(n    = 1000, 
                                    mean = 67, 
                                    sd   = 3)
                     )

my_samples |> 
  head()
```
To visualize the simulated heights, we can look at the density of the values. 

To plot simulated (emprical) data we:

+ use the `geom_histogram()` 
+ and define a local aesthetic
to calculate and plot the **density** of these values. Including the `y = after_after(density)` is very important to make the plot on the same scale as the theoretical density curve!

To plot the theoretical curve we:

+ use the `state_function()` geometry
+ specify the function as the density function `dnorm()` with the specified `mean` and `sd`.

```{r}
#| message: false
#| warning: false
#| echo: true

my_samples %>%
  ggplot(aes(x = height)) +
  geom_histogram(aes(y = after_stat(density)), 
                 binwidth = 1, fill = "grey") +
  stat_function(fun = ~ dnorm(.x, mean = 67, sd = 3), 
                col = "cornflowerblue", lwd = 2) + 
  xlim(c(55, 80))
```

# Simulating Random Samples

Taking random samples of either vectors or data frames is relatively simple in R using the functions `sample()` and `slice_sample()`

#### 📖 Required Reading: [R Programming for Data Science : 20.4](https://bookdown.org/rdpeng/rprogdatascience/simulation.html#random-sampling)

A couple of things that I want to emphasize or add from the reading:

+ Pay attention to whether your random process requires sampling *with* replacement or sampling *without* replacement.
+ You can implement a random permutation by taking a sample without replacement the same length as the provided vector:

```{r}
sample(x = 1:10, size = 10)
```
+ To get a random sample of rows from a dataframe, rather than a vector, use `slice_sample()` from dplyr. You can provide a number or samples or poportion of the rows to sample:

```{r}
penguins |> 
  slice_sample(n = 5)
```

```{r}
penguins |> 
  slice_sample(prop = .01)
```

# Efficient Simulation Iteration

Often, we want to repeat a simulation (random process) many times to see how it behaves! This involves **iteration**. Thus, to implement iterative simulations, we will want to write a **function** that implements one iteration of the simulation, and then use a `map` function to efficiently repeat that function many times and save the results.

Let's take a classic problem -- the [Monty Hall Problem](https://en.wikipedia.org/wiki/Monty_Hall_problem). A question was posed by Craig F. Whittaker in Marilyn vos Savant's "Ask Marilyn" column in Parade Magazine in 1990.[^1]

> Suppose you're on a game show, and you're given the choice of three doors: Behind one door is a car; behind the others, goats. You pick a door, say No. 1, and the host, who knows what's behind the doors, opens another door, say No. 3, which has a goat. He then says to you, "Do you want to pick door No. 2?" Is it to your advantage to switch your choice?

```{r}
monty_hall <- function(x){
  
  # randomly permute the three possibilities between 3 "doors"
  doors <- sample(c("car", "goat", "goat"),
                  3)
  
  # randomly select one door as a first choice
  first_choice <- sample(1:3, 1)
  
  # find which of the remaining doors have goats behind them
  remaining_goat_doors <- setdiff(which(doors == "goat"), first_choice)
  
  # reveal one of the goat doors 
  # (if there are still 2 doors with goats, choose randomly)
  
  revealed_door <- ifelse(length(remaining_goat_doors) > 1, 
                          sample(remaining_goat_doors, 1),
                          remaining_goat_doors)

  # have the player switch their choice (aka, choose the remaining door option)
  final_prize <- doors[setdiff(1:3, c(first_choice, revealed_door))]
  
  # return if they won the car or got a goat!
  final_prize
}
```

```{r}
monty_hall()
```

Now that we have implemented one iteration of the simulation, let's repeat it many times! The mathematical solution to the problem is that the probability of getting a car if you switch is 2/3. Let's see if our simulation shows that too.

Since the output of my function is a character, I am going to use `map_chr()`. All I want to do is repeatedly run the function many times, so I will input a vector of `1:n` where `n` is how many times I want to run the simulation as the first argument `.x` to `map_chr()`:

```{r}
sim_result <- map_chr(.x = 1:1000,
                      .f = monty_hall)
```

:::callout-note

If you don't have any additional arguments to pass to the function you provide to `map` you do not need to write the function as a lambda function. You only need to provide the function *name*. But careful! This would not work if I wrote `.f = monty_hall()`.
:::

Now let's see the proportion of times that we won a car:

```{r}
mean(sim_result == "car")
```

Pretty close to the theoretical value!

[^1]: vos Savant, Marilyn (9 September 1990). "Ask Marilyn". Parade. p. 16. Archived from the original  Retrieved 5 May 2026. [Link](https://web.archive.org/web/20130121183432/http://marilynvossavant.com/game-show-problem/)