---
title: "PA 8.1: Tidy Eval Exploration"
format: html
execute: 
  error: true
  echo: true
  message: false
  warning: false
---

[Download starter qmd file](../student-versions/practice-activities/pa8-1.qmd)

We're going to explore our functions using the `penguins` dataset from the `palmerpenguins` package. You should have this package installed previously, but if not, install it again!

```{r}
#| label: setup

library(palmerpenguins)
library(tidyverse)
```

In this practice activity, you are going to explore writing "tidy" versions of some base R table functions. 

### Base R function

Let's explore the `table()` function first.

```{r}
table(penguins$species)
```

```{r}
table(penguins$species, penguins$island)
```

**1. What does the base R `table()` function do? What does it do if you provide it with one variable? What about 2?**

### Write dplyr / tidyr code to accomplish the task

**2. Using the penguins data, write code which will:**

+ count() the number of penguins for each species and island
+ pivot the table to a wide format
+ replace NA values with 0s

```{r}
# code to create table


```


**3. Now, use that code to help create a `tidy_table()` function.**

This function should accept as inputs:

1. A dataframe
2. A variable name to be the columns of the table
3. A variable name to be the rows of the table

These variable names should be bare (no quotation marks around them). This means you will have to use one of the tidy evaluation techniques we learned!

```{r}
# tidy table function


```



**4. Test out your function!**

```{r}
#| eval: false

tidy_table(df = penguins, 
           row_var = species, 
           col_var = island)
```


**5. Does it work if only one variable is input?**

```{r}
#| eval: false

tidy_table(df = penguins, species)
```



